Super-resolution of Aeromagnetic Survey Data

### 2023-10-09 discussion.tex
Spell check and adjust chapter structure

%brings together the key findings and contributions of the research as a whole
% provide a synthesis of the work as a whole presented in the body of the thesis.
% It is important that this final chapter demonstrates the collective contribution to knowledge provided by the research described in the thesis,
% and provide overall conclusions that address the research aims.
% \setcounter{chapter}{1}

\section{Future work}
\label{sec:futurework}
The outputs of machine learning models are regarded with apprehension in geophysics practice due to the perceived “black box” of these methods \parencite{delaatAlgorithmicDecisionMakingBased2018,rudinStopExplainingBlack2019}, and the impact these models may have on subsequent interpretation.
As widely accepted by the geoscience community, all models are wrong, but some models are useful, and this extends to machine learning models.

@@

\subsection{SR impact on interpretation}
The super-resolution outputs from the methods presented in this thesis have improved structural accuracy over the low-resolution grids.
This is likely to have a positive impact on the detection of features such as lineaments and unit boundaries for geological interpretation from magnetic data.
However, despite the reduction of introduced structural inaccuracies with the use of open access training data in Cref{ch:paper2}, it is likely that some features with specific structural geometries will still be incorrectly super-resolved.
A domain expert may be able to recognise structures that are inconsistent using auxiliary information and the regional geological context, but machine processed super-resolution grids will use these data as provided.
In order to improve the adoption of SR for geological interpretation from potential field data, a rigorous study of the impact of SR on interpretation outcomes is required.
This could be a geological mapping task, such as the reinterpretation of regional geology using super-resolution upscaled regional magnetic TMI grids.
Comparison could then be made for the accuracy and benefit of a super-resolution method compared to the existing low-resolution interpretation.
However, such analysis will be difficult to perform with respect to different interpreters' variability in pattern recognition, especially in the context of their confidence in the reliability of the super-resolution product.
Previously, \textcite{sivarajahIdentifyingEffectiveInterpretation2013} reported significant variability between individuals when detecting geological targets from magnetic geophysics data.
A more reliable, repeatable and quantifiable approach would use automated pattern recognition methods developed for geophysical features.
For example, the automated lineament detection technique of \textcite{holdenAutomaticIdentificationResponses2011,holdenAutomatedAnalysisRegional2008} could be applied to low-resolution and SR grids, to compare their delineated features.

\subsection{SR model construction}
Several models were implemented throughout cref{ch:paper1, ch:paper2}, namely RDN, ESRGAN, and LTE\@.
% However, the most important choice in developing a performant super-resolution neural network for real-world aeromagnetic survey data was the choice of training data.
Cref{paper1} was prepared prior the release of a large and varied synthetic geology training dataset \parencite{jessellNoddyverseMassiveData2022}.
It was seen in cref{paper2} that this data improved model performance for a range of features by using transfer learning for pre-training.
% This transfer learning approach for training data is also found to be useful for RDN and ESRGAN\@.
Here, the use of synthetic and real-world line spacing sampled data approach used with LTE in cref{ch:paper2} are adopted for the RDN and ESRGAN models.
Cref{fig:modcomp} shows RDN and ESRGAN in predicting accurate features when upscaling \qty{320}{\m} data to \qty{80}{\m} line spacing.
It can be inferred that the challenge of super-resolution for these tasks is predominantly a question of providing suitable and sufficient training data, and is less constrained to neural network choices.

\begin{figure}
    \includegraphics[width=\linewidth]{fig/etc/TMI Comparison_23.pdf} % TODO Update for RDN, ESRGAN, LTE, ~3 hours ea.

    \caption{A comparison between all deep learning super-resolution methods presented in cref{paper1} and cref{paper2}, using the realistic low-resolution transform developed in cref{paper2}. The low-resolution input is presented in the top left, the high-resolution target in the top right. From left to right in the lower row; The RDN+ Network, the ESRGAN+ network, and the LTE network.}
    \label{modcomp}
\end{figure}

@@

\subsection{Image quality assessment for potential field grids}
Model performance is quantified using image quality assessments (IQA) such as peak signal-to-noise ratio (PSNR), structural similarity index (SSIM), feature similarity index (FSIM), or many others.
For supervised tasks, a full reference metric can be calculated between the super-resolution prediction and the high-resolution target.
In the Cref{ch:paper1,ch:paper2} work, PSNR and SSIM are used to quantify the performance of the proposed methods in transforming low-resolution grids to high resolution.
However, properties of magnetic grid data are distinct to the images for which these metrics are designed.
%   While SSIM uses structural information present in the Fourier transformed input in combination with local gradient magnitude for value accuracy,
Consistently high SSIM scores are assigned to grids that are perceptually poor matches to the target.
An investigation to identify a more appropriate IQA for potential field grids is required, which accounts for the distinct nature of these data and reflects the predicted quality of geological interpretation.
This may be benefit by automated analysis methods described above.

\subsection{Physics informed neural networks}
Potential fields are constrained by several well understood physical laws, and gridded data ideally conform to these properties for use in inversion applications and further numerical processing.
Geological interpretation is less reliant on this adherence, however \emph{a priori} information used in physics informed neural networks improve training speed and model performance \parencite{raissiPhysicsinformedNeuralNetworks2019}, and are especially useful in data-constrained domains such as geophysics.
It is possible to broaden the definition and extend the concept of informed networks to the use of generative adversarial networks in Cref{paper1}.
In these networks, the informing method is termed discriminator loss, and is performed by a classification network conditioned to identify images as `realistic' or `not realistic', depending on features learnt to be important within the target data by the discriminator.
Using crafted or learnt criterions to inform a neural network greatly assists with learning high quality models, and either approach may be useful for to geophysics in addressing the paucity of ground truth.
In addition, demonstrating the enforcement of physical properties of potential fields for model outputs may address apprehension for machine learning in geophysics practice.

@@

The challenge of training data availibility was addressed in Cref{ch:paper2}, where we investigated the use of geologically varied synthetic data to train a baseline model, fine-tuning with real-world data, and simulating realistic line sampling to create low-resolution data.
We implemented the contemporary LTE network to perform this investigation, these data sampled at four times wider line spacing showed increased structural accuracy when fine-tuned with real-world data.
Additionally, spurious features in the baseline model were correctly resolved when fine-tuned.
We identified opportunities for more detailed investigations using a filtered approach in selecting training data, and these are outlined in \Cref{sec:futurework}.

The contribution in Cref{ch:paper3} used coordinate multilayer perceptron neural networks to learn a representation of the function that describes a surveyed potential field extent.
To my knowledge, this is the first time implicit neural representation has been reported for use with aerogeophysical survey data in the literature.
These networks are highly suited for processing spatial data such as point sampled geophysics surveys.
The contribution is capable of predicting regularly spaced grids from scattered survey point data, and a novel method using the representation to calculate derivatives of a potential field, both fundamental to interpretation and processing in geophysics.
By commencing the methods from point sample data, the number of individual processing steps is reduced, and the application is fully trainable.
Coordinate MLPs for implicit representation are a currently burgeoning field of research in computer science, and we identify opportunities in extending advances in this research to future applications in geophysics, as outlined in \Cref{sec:futurework}.

>>>>

%brings together the key findings and contributions of the research as a whole
% provide a synthesis of the work as a whole presented in the body of the thesis.
% It is important that this final chapter demonstrates the collective contribution to knowledge provided by the research described in the thesis,
% and provide overall conclusions that address the research aims.
\setcounter{chapter}{4}

\chapter{Future work}
\label{ch:futurework}
The outputs of machine learning models are regarded with apprehension in geophysics practice due to the perceived “black box” of these methods \parencite{delaatAlgorithmicDecisionMakingBased2018,rudinStopExplainingBlack2019}, and the impact these models may have on subsequent interpretation.
As widely accepted by the geoscience community, all models are wrong, but some models are useful, and this extends to machine learning models.

@@

\section{SR impact on interpretation}
The super-resolution outputs from the methods presented in this thesis have improved structural accuracy over the low-resolution grids.
This is likely to have a positive impact on the detection of features such as lineaments and unit boundaries for geological interpretation from magnetic data.
However, despite the reduction of introduced structural inaccuracies with the use of open access training data in Cref{ch:paper2}, it is likely that some features with specific structural geometries will still be incorrectly super-resolved.
A domain expert may be able to recognise structures that are inconsistent using auxiliary information and the regional geological context, but machine processed super-resolution grids will use these data as provided.
In order to improve the adoption of SR for geological interpretation from potential field data, a rigorous study of the impact of SR on interpretation outcomes is required.
This could be a geological mapping task, such as the reinterpretation of regional geology using super-resolution upscaled regional magnetic TMI grids.
Comparison could then be made for the accuracy and benefit of a super-resolution method compared to the existing low-resolution interpretation.
However, such analysis will be difficult to perform with respect to different interpreters' variability in pattern recognition, especially in the context of their confidence in the reliability of the super-resolution product.
Previously, \textcite{sivarajahIdentifyingEffectiveInterpretation2013} reported significant variability between individuals when detecting geological targets from magnetic geophysics data.
A more reliable, repeatable and quantifiable approach would use automated pattern recognition methods developed for geophysical features.
For example, the automated lineament detection technique of \textcite{holdenAutomaticIdentificationResponses2011,holdenAutomatedAnalysisRegional2008} could be applied to low-resolution and SR grids, to compare their delineated features.

\section{SR model construction}
Several models were implemented throughout cref{ch:paper1, ch:paper2}, namely RDN, ESRGAN, and LTE\@.
% However, the most important choice in developing a performant super-resolution neural network for real-world aeromagnetic survey data was the choice of training data.
Cref{paper1} was prepared prior the release of a large and varied synthetic geology training dataset \parencite{jessellNoddyverseMassiveData2022}.
It was seen in cref{paper2} that this data improved model performance for a range of features by using transfer learning for pre-training.
% This transfer learning approach for training data is also found to be useful for RDN and ESRGAN\@.
Here, the use of synthetic and real-world line spacing sampled data approach used with LTE in cref{ch:paper2} are adopted for the RDN and ESRGAN models.
Cref{fig:modcomp} shows RDN and ESRGAN in predicting accurate features when upscaling \qty{320}{\m} data to \qty{80}{\m} line spacing.
It can be inferred that the challenge of super-resolution for these tasks is predominantly a question of providing suitable and sufficient training data, and is less constrained to neural network choices.

\begin{figure}
    \includegraphics[width=\linewidth]{fig/etc/TMI Comparison_23.pdf} % TODO Update for RDN, ESRGAN, LTE, ~3 hours ea.
    \caption{
        A comparison between all deep learning super-resolution methods presented in cref{paper1} and cref{paper2}, using the realistic low-resolution transform developed in cref{paper2}.
        a) The low-resolution input, b) The high-resolution target, c) The RDN+ Network, d) The ESRGAN+ network, and e) The LTE network.}
    \label{modcomp}
\end{figure}

@@

\section{Image quality assessment for potential field grids}
Model performance is quantified using image quality assessments (IQA) such as peak signal-to-noise ratio (PSNR), structural similarity index (SSIM), feature similarity index (FSIM), or many others.
For supervised tasks, a full reference metric can be calculated between the super-resolution prediction and the high-resolution target.
In the Cref{ch:paper1,ch:paper2} work, PSNR and SSIM are used to quantify the performance of the proposed methods in transforming low-resolution grids to high resolution.
However, properties of magnetic grid data are distinct to the images for which these metrics are designed.
%   While SSIM uses structural information present in the Fourier transformed input in combination with local gradient magnitude for value accuracy,
Consistently high SSIM scores are assigned to grids that are perceptually poor matches to the target.
An investigation to identify a more appropriate IQA for potential field grids is required, which accounts for the distinct nature of these data and reflects the predicted quality of geological interpretation.
This may be benefit by automated analysis methods described above.

\section{Physics informed neural networks}
Potential fields are constrained by several well understood physical laws, and gridded data ideally conform to these properties for use in inversion applications and further numerical processing.
Geological interpretation is less reliant on this adherence, however \emph{a priori} information used in physics informed neural networks improve training speed and model performance \parencite{raissiPhysicsinformedNeuralNetworks2019}, and are especially useful in data-constrained domains such as geophysics.
It is possible to broaden the definition and extend the concept of informed networks to the use of generative adversarial networks in Cref{paper1}.
In these networks, the informing method is termed discriminator loss, and is performed by a classification network conditioned to identify images as `realistic' or `not realistic', depending on features learnt to be important within the target data by the discriminator.
Using crafted or learnt criterions to inform a neural network greatly assists with learning high quality models, and either approach may be useful for to geophysics in addressing the paucity of ground truth.
In addition, demonstrating the enforcement of physical properties of potential fields for model outputs may address apprehension for machine learning in geophysics practice.

@@

The challenge of training data availability was addressed in Cref{ch:paper2}, where we investigated the use of geologically varied synthetic data to train a baseline model, fine-tuning with real-world data, and simulating realistic line sampling to create low-resolution data.
We implemented the contemporary LTE network to perform this investigation, these data sampled at four times wider line spacing showed increased structural accuracy when fine-tuned with real-world data.
Additionally, spurious features in the baseline model were correctly resolved when fine-tuned.
We identified opportunities for more detailed investigations using a filtered approach in selecting training data, and these are outlined in \Cref{ch:futurework}.

The contribution in Cref{ch:paper3} used coordinate multilayer perceptron neural networks to learn a representation of the function that describes a surveyed potential field extent.
To my knowledge, this is the first time implicit neural representation has been reported for use with aerogeophysical survey data in the literature.
These networks are highly suited for processing spatial data such as point sampled geophysics surveys.
The contribution is capable of predicting regularly spaced grids from scattered survey point data, and a novel method using the representation to calculate derivatives of a potential field, both fundamental to interpretation and processing in geophysics.
By commencing the methods from point sample data, the number of individual processing steps is reduced, and the application is fully trainable.
Coordinate MLPs for implicit representation are a currently burgeoning field of research in computer science, and we identify opportunities in extending advances in this research to future applications in geophysics, as outlined in \Cref{ch:futurework}.

\printbibliography{}

### 2023-10-09 intro.tex
Spell check and adjust chapter structure

\documentclass[manuscript.tex]{subfiles}
\begin{document}
\setcounter{chapter}{1}

% Ensure acronyms correctly used

The subsurface continuation of exposed geology can be extrapolated and interpolated using structures mapped in surveys of the Earth's naturally occuring potential fields.
Potential field surveys offer low-cost investigations over a large spatial extent, and routinely used for geological mapping and interpretation \parencite{nabighian75thAnniversaryHistorical2005}
For this reason, magnetic field surveys are commonly used by geophysicists for geological mapping and interpretation toward understanding the Earth in three dimensions.
Scattered surveys are widely used when interpolated to grid rasters, and interpreted individually or integrated with many other datasets for predicting the subsurface for processes such as numerical modelling and inversion.
In these tasks, the spatial resolution of the grid is a critical factor in their ability to define geological detail \parencite{islesGeologicalInterpretationAeromagnetic2018}.
The role of high-resolution details in these geophysical methods is becoming more important to exploration as resources are increasingly sought from below sedimentary cover.
Resolution enhancement has seen recent success from the use of deep learning \parencite{moserHitchhikerGuideSuperResolution2023}, however it is not understood if these methods could extend to geophysics data, which are highly distinct from natural images.

@@

The above aims are achieved using case studies with open access aeromagnetic data provided by Geoscience Australia.

@@

\subsection{Survey data regularisation}
\label{sec:introgrids}
Survey data can be interpreted as one dimensional transects, but more commonly they are regularised to a two-dimensional grid raster for interpretation in a process termed gridding.
The resulting interpolated array contains uniformly spaced nodes and is referred to as a grid.
Grids are raster data, where each pixel stores a scalar measurement of a specific property.
In geophysics and throughout this thesis, these are referred to as cells, and the property is a measure of the strength of the magnetic field at a specific location.
The field strength measured is the Total Magnetic Intensity (TMI), which is the vector sum of each directional vector of the potential \parencite{blakelyPotentialTheoryGravity1996}.
This scalar value is recorded in nanotesla and may be negative with an unbounded range that can exceed \qty{50,000}{\nano\tesla} in the Earth's environment.
The location recorded at a sample point is the geographic latitude and longitude, and altitude, acquired with high-precision differential GPS, alongside radar altimeter elevation.
These data may be transformed to a flat projected system, where spatial dimensions are expressed in meters, used throughout this thesis.
There are numerous conventional gridding methods, and novel methods are frequently proposed in the literature.
The simplest use surrounding samples for interpolation, such as choosing the closest sample value or applying a transformation function to the neighbouring samples.
Of those most commonly used in potential field geophysics, these are minimum curvature \parencite{briggsMachineContouringUsing1974}, splines \parencite{bhattacharyyaBicubicSplineInterpolation1969,shureHarmonicSplinesGeomagnetic1982,smithGriddingContinuousCurvature1990}, and equivalent sources \parencite{dampneyEquivalentSourceTechnique1969, solerBetterStrategyInterpolating2020}.
Splines are used in bi-directional gridding, which is well-suited for aeromagnetic surveys, where the sample data are first interpolated in the line parallel direction, and the intermediate product is interpolated in the line perpendicular direction \parencite{dentithGeophysicsMineralExploration2014}.
A method known as kriging is used in other areas of geoscience, such as geochemical mapping, and makes use of statistical properties of the known distribution \parencite{hansenInterpretiveGriddingAnisotropic1993,davis1986statistics}.
Some recently proposed gridding methods include the methods of \textcite{naprstekNewMethodInterpolating2019}, \textcite{xuGravityAnomalyReconstruction2019}, or \textcite{chenPotentialFieldData2022}.
Each interpolation method attempts to predict regularly spaced grids by leveraging properties intrinsic to the data, with the possible explicit integration of \emph{a priori} knowledge such as geophysical laws or geological context.
A review of spatial interpolation methods is provided in \textcite{liReviewComparativeStudies2011}.

@@

Geoscience data faces a number of challenges distinct from natural images. These have been outlined in \textcite{karpatneMachineLearningGeosciences2019}, and relevant to this work are:
\begin{itemize}
    \item{} Spatio-Temporal Structure: Aeromagnetic grids are spatially corellated, neighbouring cells contain similar values arising from the same causative volume.
    \item{} Multi-Source Multi-Resolution Data: One location may be covered by mutiple aeromagnetic surveys with different resolutions, or the same survey may be presented with different levels of processing.
    \item{} Poor Quality of Data: Sample data are subject to numerous sources of noise, which combine to become a significant fraction of the gridded data value, or be present as aliased frequencies.
    \item{} Small Sample Size: While total aeromagnetic coverage in Australia is high, the high-resolution extents are spatially limited and biased by geological domain.
\end{itemize}

@@

\subsection{Deep learning context}
\label{sec:introml}
Machine learning (ML) is a very broad family of techniques for learning functions from data.
Artificial neural networks are a subset of techniques in ML, inspired by a model of the biological neuron \parencite{bishopNeuralNetworksPattern1995}.
It has been established that if sufficient conditions are met, neural networks can approximate any function \parencite{hornikMultilayerFeedforwardNetworks1989}.
A mathematical model of a neuron forms the basis of these networks, and the number of neurons within the network and their arrangement are one primary descriptor of a neural network. 
Neurons are paremeterised by a multiplicative weight term, an added bias term, and an activation function to scale the result.
Each neuron applies these operations to the input, which may either be the input data, or the output of earlier neurons.
The objective of training a neural network is to set the values for these for each neuron in the network, such that the trained network (or model) generalises to novel data.
This is performed by back propogation \parencite{rumelhartLearningRepresentationsBackpropagating1988}, a technique to navigate toward the lowest error solution by calculating the rate of improvement resulting from parameter updates.
A practical introduction to the practice of training machine learning is available from \textcite{stevensDeepLearningPyTorch2020}, while a more formal introduction is provided in \textcite{bishopPatternRecognitionMachine2006}.
Many activation functions have been proposed, with the rectified linear unit (ReLU) being used in most tasks throughout this thesis.
The activation function performs scaling but in many classes of activation function it also performs the important task of scaling the result to zero, which can be interpreted as the lack of a specific feature being present \parencite{williamsLogicActivationFunctions1986}.
However, recent work on periodic activation functions such as the sinusoid \parencite{sitzmann2019siren} enable the networks used in Cref{ch:paper3}.
The other primary descriptor of a neural network is its architecture, refering to the configuration of its intermediate layers between the input and output.
These architectures are described below.

@@

Deep learning is a field of contemporary machine learning, where multiple layers of feature learning are undertaken between the input to a method and the output.
The current generation of deep learning is reported in review articles to have begun in 2006 \parencite{dengDeepLearningMethods2014,goodfellowDeepLearning2016a}, and has pervaded many industries and research fields.
These review articles contain thorough background on the implementation and practice of various architectures for undertaking deep learing, as well as earlier machine learning or numerical models that can address tasks now approached with deep learning.
The deep learning tasks explored in this thesis are super-resolution (SR) for the objective of upsampling resolution enhancement, and implicit neural representation (INR).
The network architectures implemented in this thesis to address the tasks include convolutional neural networks (CNN), generative adversarial networks (GAN), and multilayer perceptrons (MLP).
A brief introduction to these topics is provided here.

>>>>

\documentclass[manuscript.tex]{subfiles}
\setcounter{chapter}{0}
\begin{document}

\chapter{Introduction}
\label{ch:intro}
The subsurface continuation of exposed geology can be extrapolated and interpolated using structures mapped in surveys of the Earth's naturally occuring potential fields.
Potential field surveys offer low-cost investigations over a large spatial extent, and routinely used for geological mapping and interpretation \parencite{nabighian75thAnniversaryHistorical2005}
For this reason, magnetic field surveys are commonly used by geophysicists for geological mapping and interpretation toward understanding the Earth in three dimensions.
Scattered surveys are widely used when interpolated to grid rasters, and interpreted individually or integrated with many other datasets for predicting the subsurface for processes such as numerical modelling and inversion.
In these tasks, the spatial resolution of the grid is a critical factor in their ability to define geological detail \parencite{islesGeologicalInterpretationAeromagnetic2018}.
The role of high-resolution details in these geophysical methods is becoming more important to exploration as resources are increasingly sought from below sedimentary cover.
Resolution enhancement has seen recent success from the use of deep learning \parencite{moserHitchhikerGuideSuperResolution2023}, however it is not understood if these methods could extend to geophysics data, which are highly distinct from natural images.

@@

The above aims are investigated using case studies with open access aeromagnetic data provided by Geoscience Australia.

@@

\subsection{Survey data regularisation}
\label{sec:introgrids}
Survey data can be interpreted as one dimensional transects, but more commonly they are regularised to a two-dimensional grid raster for interpretation in a process termed gridding.
The resulting interpolated array contains uniformly spaced nodes and is referred to as a grid.
Grids are raster data, where each pixel stores a scalar measurement of a specific property.
In geophysics and throughout this thesis, these are referred to as cells, and the property is a measure of the strength of the magnetic field at a specific location.
The field strength measured is the Total Magnetic Intensity (TMI), which is the vector sum of each directional vector of the potential \parencite{blakelyPotentialTheoryGravity1996}.
This scalar value is recorded in nanotesla and may be negative with an unbounded range that can exceed \qty{50,000}{\nano\tesla} in the Earth's environment.
The location recorded at a sample point is the geographic latitude and longitude, and altitude, acquired with high-precision differential GPS, alongside radar altimeter elevation.
These data may be transformed to a flat projected system, where spatial dimensions are expressed in metres, used throughout this thesis.
There are numerous conventional gridding methods, and novel methods are frequently proposed in the literature.
The simplest use surrounding samples for interpolation, such as choosing the closest sample value or applying a transformation function to the neighbouring samples.
Of those most commonly used in potential field geophysics, these are minimum curvature \parencite{briggsMachineContouringUsing1974}, splines \parencite{bhattacharyyaBicubicSplineInterpolation1969,shureHarmonicSplinesGeomagnetic1982,smithGriddingContinuousCurvature1990}, and equivalent sources \parencite{dampneyEquivalentSourceTechnique1969, solerBetterStrategyInterpolating2020}.
Splines are used in bi-directional gridding, which is well-suited for aeromagnetic surveys, where the sample data are first interpolated in the line parallel direction, and the intermediate product is interpolated in the line perpendicular direction \parencite{dentithGeophysicsMineralExploration2014}.
A method known as kriging is used in other areas of geoscience, such as geochemical mapping, and makes use of statistical properties of the known distribution \parencite{hansenInterpretiveGriddingAnisotropic1993,davis1986statistics}.
Some recently proposed gridding methods include the methods of \textcite{naprstekNewMethodInterpolating2019}, \textcite{xuGravityAnomalyReconstruction2019}, or \textcite{chenPotentialFieldData2022}.
Each interpolation method attempts to predict regularly spaced grids by leveraging properties intrinsic to the data, with the possible explicit integration of \emph{a priori} knowledge such as geophysical laws or geological context.
A review of spatial interpolation methods is provided in \textcite{liReviewComparativeStudies2011}.

@@

Geoscience data faces a number of challenges distinct from natural images. These have been outlined in \textcite{karpatneMachineLearningGeosciences2019}, and relevant to this work are:
\begin{itemize}
    \item{} Spatio-Temporal Structure: Aeromagnetic grids are spatially corellated, neighbouring cells contain similar values arising from the same causative volume.
    \item{} Multi-Source Multi-Resolution Data: One location may be covered by multiple aeromagnetic surveys with different resolutions, or the same survey may be presented with different levels of processing.
    \item{} Poor Quality of Data: Sample data are subject to numerous sources of noise, which combine to become a significant fraction of the gridded data value, or be present as aliased frequencies.
    \item{} Small Sample Size: While total aeromagnetic coverage in Australia is high, the high-resolution extents are spatially limited and biased by geological domain.
\end{itemize}

@@

\subsection{Deep learning context}
\label{sec:introml}
Machine learning (ML) is a very broad family of techniques for learning functions from data.
Artificial neural networks are a subset of techniques in ML, inspired by a model of the biological neuron \parencite{bishopNeuralNetworksPattern1995}.
It has been established that if sufficient conditions are met, neural networks can approximate any function \parencite{hornikMultilayerFeedforwardNetworks1989}.
A mathematical model of a neuron forms the basis of these networks, and the number of neurons within the network and their arrangement are one primary descriptor of a neural network. 
Neurons are paremeterised by a multiplicative weight term, an added bias term, and an activation function to scale the result.
Each neuron applies these operations to the input, which may either be the input data, or the output of earlier neurons.
The objective of training a neural network is to set the values for these for each neuron in the network, such that the trained network (or model) generalises to novel data.
This is performed by back propagation \parencite{rumelhartLearningRepresentationsBackpropagating1988}, a technique to navigate toward the lowest error solution by calculating the rate of improvement resulting from parameter updates.
A practical introduction to the practice of training machine learning is available from \textcite{stevensDeepLearningPyTorch2020}, while a more formal introduction is provided in \textcite{bishopPatternRecognitionMachine2006}.
Many activation functions have been proposed, with the rectified linear unit (ReLU) being used in most tasks throughout this thesis.
The activation function performs scaling but in many classes of activation function it also performs the important task of scaling the result to zero, which can be interpreted as the lack of a specific feature being present \parencite{williamsLogicActivationFunctions1986}.
However, recent work on periodic activation functions such as the sinusoid \parencite{sitzmann2019siren} enable the networks used in Cref{ch:paper3}.
The other primary descriptor of a neural network is its architecture, referring to the configuration of its intermediate layers between the input and output.
These architectures are described below.

@@

Deep learning is a field of contemporary machine learning, where multiple layers of feature learning are undertaken between the input to a method and the output.
The current generation of deep learning is reported by review articles to have begun in 2006 \parencite{dengDeepLearningMethods2014,goodfellowDeepLearning2016a}, and has pervaded many industries and research fields.
These review articles contain thorough background on the implementation and practice of various architectures for undertaking deep learing, as well as earlier machine learning or numerical models that can address tasks now approached with deep learning.
The deep learning tasks explored in this thesis are super-resolution (SR) for the objective of upsampling resolution enhancement, and implicit neural representation (INR).
The network architectures implemented in this thesis to address the tasks include convolutional neural networks (CNN), generative adversarial networks (GAN), and multilayer perceptrons (MLP).
A brief introduction to these topics is provided here.

### 2023-10-09 paper2.tex
Spell check and adjust chapter structure

\documentclass[manuscript.tex]{subfiles}
\begin{document}
% \setcounter{chapter}{1}
\title{Super-resolution of Aeromagnetic Survey Data}
\author[1*]{Luke Thomas Smith}
\author[1]{Tom Horrocks}
\author[1]{Eun-Jung Holden}
\author[1]{Daniel Wedge}
\author[2]{Naveed Akhtar}
\affil{Centre for Data-Driven Geoscience, School of Earth Sciences, The University of Western Australia (M006), 35 Stirling Highway, 6009 Perth, Australia}
\affil{Department of Computer Science and Software Engineering, The University of Western Australia (M002), 35 Stirling Highway, 6009 Perth, Australia}
\date{\today}
\maketitle{}

@@

\subsection{Resolution in geophysical data}
\label{sec:resinsurveys}
In exploration geophysics, the Earth's naturally occurring magnetic or gravitational potential fields can be used to survey the petrophysical expression of underlying geology.
These can be interpreted to inform geological mapping, resource modelling, and enhance the understanding of the subsurface.
Depending on the scale of investigation, surveys may be carried out from ground or airborne sensors.
Detailed ground-based surveys can be sampled at point spacings of tens of metres while avoiding local obstacles, but at a regional scale, aircraft are used to provide rapid and uniform coverage regardless of access constraints.
These airborne surveys collect data at wide line spacing to minimise cost, with lines in Western Australian low-resolution regional surveys typically \qty{400}{\m} apart, however, open access high-resolution surveys with line spacings of \qty{100}{\m} or closer are increasingly common.
Sampling along line is typically performed at \qty{10}{\Hz} or faster, which corresponds to several meters between samples at a nominal flight speed of \qty{70}{\m\per\s} \parencite{goodwinAirborneMagneticRadiometric2023}.
Because the power of shorter wavelengths in potential fields diminish with increasing distances from the source, aeromagnetic flights are flown as low as safety and topography permits.
Despite the use of precise navigation systems and skilled operators, these data are not collected in perfectly uniform lines at constant height, and the samples also contain noise components.

@@

\section{Discussion}
\label{sec:2discussion}
The Noddyverse synthetic training dataset comprises geology models with a variety of realistic petrophysics and geological histories, and was sampled using simple but realistic airborne survey methodology.
Furthermore, real survey data used to fine tune the model were extracted from extents across a range of different terranes in Western Australia.
As such, the method and trained models generalise to the geological terranes represented in the Noddyverse and high-resolution state grid extents.
Granitic textures in real grid data are not represented in the synthetic Noddyverse training set, and show the least improvement by the method.
Different geological units present different features in magnetic potential field grids, with deformed igneous units typically causing abundant contextually important short-wavelength content.
These features, such as faulted or folded dykes, tend to show the most perceptual improvement in their structure following super-resolution upscaling.
This is due to the high performance of the method for reducing aliasing effects from the gridding process, and is apparent in the accurate reconstruction of linear features from discrete “string of beads” aliased features.
Long-wavelength features and smooth gradients, such as those present in sedimentary units or areas of cover are accurately upscaled in the proposed method, however these are also adequately upscaled with simple cubic spline interpolation of the low-resolution grid because they do not require predicted high frequency components.
The exception for super-resolution having higher quality is in areas of low magnitude TMI of only a few nanoTesla across the entire extent.
In these cases, the otherwise minor noise introduced by the super-resolution process dominates the residual.

>>>>

\documentclass[manuscript.tex]{subfiles}
\begin{document}
\setcounter{chapter}{2}
\title{Super-resolution of Aeromagnetic Survey Data}
\author[1*]{Luke Thomas Smith}
\author[1]{Tom Horrocks}
\author[1]{Eun-Jung Holden}
\author[1]{Daniel Wedge}
\author[2]{Naveed Akhtar}
\affil{Centre for Data-Driven Geoscience, School of Earth Sciences, The University of Western Australia (M006), 35 Stirling Highway, 6009 Perth, Australia}
\affil{Department of Computer Science and Software Engineering, The University of Western Australia (M002), 35 Stirling Highway, 6009 Perth, Australia}
\date{\today}
\maketitle{}

@@

\subsection{Resolution in geophysical data}
\label{sec:resinsurveys}
In exploration geophysics, the Earth's naturally occurring magnetic or gravitational potential fields can be used to survey the petrophysical expression of underlying geology.
These can be interpreted to inform geological mapping, resource modelling, and enhance the understanding of the subsurface.
Depending on the scale of investigation, surveys may be carried out from ground or airborne sensors.
Detailed ground-based surveys can be sampled at point spacings of tens of metres while avoiding local obstacles, but at a regional scale, aircraft are used to provide rapid and uniform coverage regardless of access constraints.
These airborne surveys collect data at wide line spacing to minimise cost, with lines in Western Australian low-resolution regional surveys typically \qty{400}{\m} apart, however, open access high-resolution surveys with line spacings of \qty{100}{\m} or closer are increasingly common.
Sampling along line is typically performed at \qty{10}{\Hz} or faster, which corresponds to several metres between samples at a nominal flight speed of \qty{70}{\m\per\s} \parencite{goodwinAirborneMagneticRadiometric2023}.
Because the power of shorter wavelengths in potential fields diminish with increasing distances from the source, aeromagnetic flights are flown as low as safety and topography permits.
Despite the use of precise navigation systems and skilled operators, these data are not collected in perfectly uniform lines at constant height, and the samples also contain noise components.

@@

\section{Discussion}
\label{sec:2discussion}
The Noddyverse synthetic training dataset comprises geology models with a variety of realistic petrophysics and geological histories, and was sampled using simple but realistic airborne survey methodology.
Furthermore, real survey data used to fine tune the model were extracted from extents across a range of different terranes in Western Australia.
As such, the method and trained models generalise to the geological terranes represented in the Noddyverse and high-resolution state grid extents.
Granitic textures in real grid data are not represented in the synthetic Noddyverse training set, and show the least improvement by the method.
Different geological units present different features in magnetic potential field grids, with deformed igneous units typically causing abundant contextually important short-wavelength content.
These features, such as faulted or folded dykes, tend to show the most perceptual improvement in their structure following super-resolution upscaling.
This is due to the high performance of the method for reducing aliasing effects from the gridding process, and is apparent in the accurate reconstruction of linear features from discrete “string of beads” aliased features.
Long-wavelength features and smooth gradients, such as those present in sedimentary units or areas of cover are accurately upscaled in the proposed method, however these are also adequately upscaled with simple cubic spline interpolation of the low-resolution grid because they do not require predicted high frequency components.
The exception for super-resolution having higher quality is in areas of low magnitude TMI of only a few nanotesla across the entire extent.
In these cases, the otherwise minor noise introduced by the super-resolution process dominates the residual.
